## Outline

1. What is survival analysis?
    
1. Visualizing Kaplan-Meier

1. Visualizing Competing Risks

1. Advanced Topics in {ggsurvfit}

    - Combining figures
    - How Risk Tables are constructed
    - CDISC Helpers

## Survival Analysis

This is the briefest overview. 
For more details, review the resources below.

::: {.small}

> Clark, T., Bradburn, M., Love, S., & Altman, D. (2003). _Survival analysis part I: Basic concepts and first analyses_. 232-238. ISSN 0007-0920.

> M J Bradburn, T G Clark, S B Love, & D G Altman. (2003). _Survival Analysis Part II: Multivariate data analysis – an introduction to concepts and methods. British Journal of Cancer, 89(3), 431-436.

> Bradburn, M., Clark, T., Love, S., & Altman, D. (2003). Survival analysis Part III: Multivariate data analysis – choosing a model and assessing its adequacy and fit_. 89(4), 605-11.

> Clark, T., Bradburn, M., Love, S., & Altman, D. (2003). _Survival analysis part IV: Further concepts and methods in survival analysis_. 781-786. ISSN 0007-0920.

:::

## Survival Analysis

> Survival times are data that measure follow-up time from a defined starting point to the occurrence of a given event, for example the time from the beginning to the end of a remission period or the time from the diagnosis of a disease to death. Standard statistical techniques cannot usually be applied because the data are often 'censored'. A survival time is described as censored when there is a follow-up time but the event has not yet occurred.

Bewick V, Cheek L, Ball J. *Statistics review 12: survival analysis*. Crit Care. 2004 Oct;8(5):389-94.


::: {.notes}
Read this aloud, then rephrase and give examples
:::

## Survival Analysis

:::: {.columns}

::: {.column width="50%"}

Because survival data are common in many fields, it also goes by other names.

**Other Names**

- Reliability analysis

- Duration analysis

- Event history analysis

- Time-to-event analysis

:::
::: {.column width="50%"}

**Examples**

- Time from surgery to death

- Time from start of treatment to progression

- Time from HIV infection to development of AIDS

- Time to heart attack

- Time to onset of substance abuse

- Time to initiation of sexual activity
 
- Time to machine malfunction

:::
::::

## Survival Analysis, No Censoring

:::: {.columns}

::: {.column width="60%"}

```{r echo=FALSE, fig.width=6}
set.seed(123)

df_censoring <-
  tibble(full_time = rgamma(n = 5, shape = 1)) |> 
  mutate(
  obs_time = ifelse(
    row_number() < 3, 
    full_time,
    full_time * runif(n = 5)
  ),
  death = obs_time == full_time,
  id = rev(row_number())
)

df_censoring |> 
  ggplot(aes(y = id)) +
  geom_segment(
    aes(yend = id, xend = 0, x = full_time),
    linetype = "solid",
    linewidth = 8,
    lineend='round',
    color = "#3c523c"
  ) +
  emoGG::geom_emoji(
    aes(y = id, x = full_time),
    emoji = "1f480"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold"),
        axis.title.y = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        aspect.ratio = 1) +
  labs(
    title = "Time from Treatment to Death",
    x = "Time"
  )
```

:::

::: {.column width="40%"}
- The skulls are the time when each of these patients die

- We observe the follow-up in the solid line

- How can we summarize the time from treatment to death?

:::

::::


::: {.notes}
- We can use a simple mean, median, etc.
:::

## Survival Analysis, Censoring

:::: {.columns}

::: {.column width="60%"}

```{r echo=FALSE, fig.width=6}
df_censoring |> 
  ggplot(aes(y = id)) +
  geom_segment(
    aes(yend = id, xend = full_time, x = obs_time),
    linetype = "dotted",
    linewidth = 2,
    color = "grey",
    lineend='round'
  ) +
  geom_segment(
    aes(yend = id, xend = 0, x = obs_time),
    linetype = "solid",
    linewidth = 8,
    lineend='round',
    color = "#3c523c"
  ) +
  emoGG::geom_emoji(
    aes(y = id, x = full_time),
    emoji = "1f480"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold"),
        axis.title.y = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        aspect.ratio = 1) +
  labs(
    title = "Time from Treatment to Death",
    x = "Time"
  )
```

:::

::: {.column width="40%"}
- The skulls are the time when each of these patients die

- We observe the follow-up in the solid line

- The dashed line in __unobserved__

- How can we summarize the average time from treatment to death?

:::

::::

::: {.notes}
Censoring occurs if a subject has not experienced the event of interest by the end of data collection.

A subject may be censored due to:

- Loss to follow-up

- Withdrawal from study

- No event by end of fixed study period

When censoring is present, the analytic method MUST take this into account.
:::

## Survival Analysis Methods


```{r echo=FALSE}
tribble(
  ~method, ~survival_method,
  "Means/Median/Percentiles", "Kaplan-Meier Estimator",
  "t-test/Rank-sum Test", "Log-rank Test",
  "Linear/Logistic Regression", "Cox Proportional Hazards Regression"
) |> 
  gt::gt() |> 
  gt::tab_options(table.font.size = 34) |> 
  gt::cols_label(
    method = gt::md("**Standard Method**"),
    survival_method = gt::md("**Survival Method**")
  ) |> 
  gt::tab_footnote(
    "There are multiple methods available tailored to specific cases and these are the most common.",
    locations = gt::cells_column_labels(columns = survival_method)
  ) |> 
  gt::tab_footnote(
    "We will not be covering regression methods today.",
    locations = gt::cells_body(columns = survival_method, rows = str_detect(survival_method, "Cox"))
  )
```


::: {.notes}
We'll primarily be covering Kaplan-Meier methods today.
:::


