# Additional examples

# Default and styled modifications with ggplot2

```{r}
#| echo: false
#| message: false
library(ggsurvfit)
library(tidyverse)
```

## {ggsurvfit} defaults 

```{r}
gg_default <-
  survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit(size = 1) +
  add_confidence_interval() +
  labs(title = "Default")

gg_default
```

## {ggplot2} styled 


```{r}
#| output-location: slide
#| code-line-numbers: "|2|3|4-8|9|10-11|12-13|14|15-18|"

gg_styled <-
  gg_default +
  coord_cartesian(xlim = c(0, 8)) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent, 
    expand = c(0.01, 0)
  ) +
  scale_x_continuous(breaks = 0:9, expand = c(0.02, 0)) +
  scale_color_manual(values = c('#54738E', '#82AC7C')) +
  scale_fill_manual(values = c('#54738E', '#82AC7C')) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol = 1)) +
  labs(
    title = "{ggplot2} styled",
    y = "Percentage Survival"
  )

gg_styled
```



# Risk tables 

## {ggsurvfit} defaults

```{r}
survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit(size = 1) +
  add_risktable()
```


## Group by statistic or strata 

```{r}
ggrisktable <-
  survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit() +
  scale_ggsurvfit() +
  add_risktable(risktable_group = "risktable_stats") 
ggrisktable
```

## Color encoding strata
 
```{r}
ggrisktable +
  add_risktable_strata_symbol(symbol = "\U25CF", size = 12)
```



## Customizing the risktable statistics

```{r}
survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit() +
  add_risktable(risktable_stats = "{n.risk} ({cum.event})") 
```



# Quantiles 

## Median summary 

```{r}
survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit(linewidth = 0.8) +
  add_censor_mark() +
  add_quantile(y_value = 0.5) +
  scale_ggsurvfit()
```

## At a given timepoint

```{r}
survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit(linewidth = 0.8) +
  add_censor_mark() +
  add_quantile(x_value = 5, linetype = "solid", linewidth = 1.0, alpha = 0.3) +
  scale_ggsurvfit()
```

## Underyling {ggsurvfit} functions 

In addition to using additional {ggplot2} functions, it is helpful to understand which underlying functions are used to create the figures.

Additional arguments can be passed to the underlying functions.

:::: {.columns}
::: {.column width="50%"}

```{r echo=FALSE}
tribble(
  ~ggsurvfit, ~underlying,
  "`ggsurvfit(...)`", "`geom_step(...)`",
  "`add_confidence_interval(...)`", "`geom_ribbon(...)`",
  "`add_censor_mark(...)`", "`geom_point(...)`",
  "`add_quantile(...)`", "`geom_segment(...)`"
) |> 
  gt::gt() |> 
  gt::cols_label(ggsurvfit = "{ggsurvfit}",
                 underlying = "Underlying {ggplot2}") |> 
  gt::fmt_markdown()
```


```{r}
#| code-line-numbers: "|4|5-7|8-9"

p <- 
  survfit2(Surv(time, status) ~ 1, 
           data = df_colon) |> 
  ggsurvfit(color = "#508050") +
  add_confidence_interval(
    fill = "#508050"
  ) +
  add_censor_mark(color = "#508050", 
                  alpha = 100)
```

:::

::: {.column width="50%"}

```{r echo=FALSE}
p
```

:::

::::

## Further Risktable Customization 

- Here, we customize the risktable portion of the plot.

- We can style the plot area by passing ggplot2 functions to `add_risktable(theme)` argument.


```{r}
#| output-location: slide
survfit2(Surv(time, status) ~ surg, data = df_colon) %>%
  ggsurvfit(linewidth = 1) +
  add_risktable(
    risktable_height = 0.33,
    size = 4, # increase font size of risk table statistics
    theme =   # increase font size of risk table title and y-axis label
      list(
        theme_risktable_default(axis.text.y.size = 11, 
                                plot.title.size = 11),
        theme(plot.title = element_text(face = "bold"))
      )
  )
```




# Another Example

```{r}
#| output-location: slide
#| code-line-numbers: "|2|3|4-5|6|7|8|10-12"
survfit2(Surv(time, status) ~ surg, data = df_colon) |>
  ggsurvfit(type = "risk", linewidth = 1.2) +
  add_confidence_interval() +
  add_risktable(risktable_stats = "n.risk") +
  add_risktable_strata_symbol(symbol = "\U25CF", size = 17) +
  add_quantile(x_value = 5, linetype = "dotted", linewidth = 0.8) +
  add_censor_mark(size = 2, alpha = 0.2) +
  add_pvalue(caption = "Log-rank {p.value}") +
  scale_ggsurvfit() +
  scale_color_manual(values = c('#54738E', '#82AC7C')) +
  scale_fill_manual(values = c('#54738E', '#82AC7C')) +
  labs(y = "Risk of Recurrence")
```
